# Experiment setup {#sec-experiment}


## Data {#sec-data}

For this study, we use a dataset spanning five years (Dec. 2017- July 2022) of consumption data obtained from EPSS. Rigorous checks were conducted to ensure the consistency and completeness of the collected data. From the extensive pool of pharmaceutical products within EPSS, we strategically selected a set of 33 key pharmaceuticals, encompassing various programs and representing different classes of drugs. This deliberate selection enriches the research by providing a comprehensive representation across the pharmaceutical landscape.

We are presenting the graph for some of the pharmaceuticals sales for the last five years by using visualization for the readers to undesrtand the sales of the pharmaceuticals.

### Explaoratory Data Analysis with consumption data

Given the number of produts in this study, we created several data plot and computed features of the time series, including the strength of trend and seasonality to better understadn data. @fig-feature shows the strength of trend versus seasonality. Each point represents one time series with the strength of trend in x-axis and the strength of seasonality in y-axis. Both measures are on a scale of [0,1]. the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by @mstl.


```{r}
#| label: fig-feature
#| cache: true
#| out.width: "70%"
#| fig.align: center
#| fig-cap: "The strength of the trend and seasonality in the time series of pharmaceutical product consumption. The scatter plot shows 33 data points, with each point corresponding to a product."
read_csv('data/predicotrs_final_tssible_July_04_2024.csv') |> 
  janitor::clean_names()
med_qty_tsb <- med_qty |> 
  mutate(month = yearmonth(month)) |> 
  as_tsibble(index = month, key = item)

med_qty_tsb %>%
  features(quantity, feat_stl) %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_week)) +
  geom_point(alpha = 0.25) +
  labs(x = "Strength of trend", y = "Strength of weekly seasonality")
```


It is clear that there are some series showing strong trends and/or seasonality, corresponding to series at the higher levels of the hierarchy. The majority of series show low trend and seasonality. These are time series belonging to the bottom series, series related to the nature of incidents for a given control, health board and priority level. Bottom series are dominated by noise with little or no systematic patterns.

In addition to displaying the strength of trend and seasonality, we also visualize all time series, revealing various patterns such as trends, seasonality, and instances of consumption with chaotic behavior during certain months. For instance, some series exhibit both seasonality and trend, while others show a low volume of attended incidents and higher entropy, making them more volatile and challenging to forecast. At the incident level, combined with categories from other levels, many series contain zeros or low counts. This highlights the diversity of daily time series patterns in the data set. Figure @fig-dataviz2 illustrates the time plots for a few selected products.

```{r}
#| label: fig-dataviz2
#| cache: true
#| dependson: "fig-feature"
#| out.width: "100%"
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Monthly time plot of consumption. X-axis shows the month, consisting of 60 data points (months) and y-axis shows the comsumption. The panels display data from four products to give a glimpse of the consumption patterns."

(my_plot[[1]] | my_plot[[1]]) /
(my_plot[[1]] | my_plot[[1]])

```

Visualizing the data revealed that various events significantly impacted the consumption of different products, but these were not reflected in the available data. Expert insights were crucial for understanding the nature and timing of these events, filling gaps in the system, and incorporating qualitative factors that EPSS logistics system miss. Therefore, we conducted interviews with domain experts to collect information on external factors that influence consumption. These interviews allowed us to account for unusual patterns and customize our forecasting models to more accurately reflect the unique consumption behaviors of each product.


### Domain knowledge colelction though interview with EPSS experts

To gather domain knowledge, we contacted experts from the Ethiopian Pharmaceuticals Supply Service (EPSS) who work in the distribution unit. Five experts with over 10 years of experience in managing the distribution of the pharmaceuticals in EPSS agreed to participate. The researchers, along with these experts, examined the electronic bin cards of selected pharmaceuticals. A bin card is a physical or electronic record used to track the inventory levels of individual items in a storage location, such as a warehouse or stockroom. The term 'bin' refers to the specific storage location for an item. Information typically recorded on a bin card includes the Item ID, unit of measure, product description, account details, date, references, beginning balance, manufacturer, batch number, expiry date, quantity received, quantity issued, and balance. In summary, a bin card is a fundamental tool in inventory management, especially in environments where physical stock tracking is essential. It provides a straightforward and immediate way to monitor item quantities, helping to ensure effective inventory management.


We meticulously traced sales data for the past five years, triangulating graphs with bin card records to identify predictors influencing demand for each product. The expert uses their domain knowledge and experience to identify about the distribution charachterstics of the specific products and sometimes when they assume that they do not have enough history or do not remember about that specific products they call to other experts who remember specific issues and decide whetrere the specific predictos affect the distribution and if all of them agree the include it as predcitors. The data collection process took more than a month. We downloaded and analyzed over 100 bin cards per product. Dates on the bin cards were converted from the Ethiopian calendar to the Gregorian calendar for accuracy. Transactions due to storage or warehouse changes within EPSS were excluded. After cleaning the data,and discussing with the experts  we identified the following key predictors affecting distribution and sales: stock replenishment, stock outs, COVID-19, fiscal year counting, conflict, and malaria seasonality.


- Stock replenishment:  refers to the process of restocking or refilling inventory to ensure that there are sufficient quantities of products or materials available to meet demand. Whenever there was stock replenishment at the central EPSS, sales and distribution to hubs and health facilities increased. This increase was attributed to the need to restock depleted inventories and the push from central EPSS to manage space constraints. 

- Stockouts:  occur when an inventory item is completely depleted, meaning there is no available stock to meet demand. This can happen when demand exceeds the available supply, or when there is a delay or failure in the replenishment process. Resulted in no transactions and a decline in sales. 

- COVID-19: COVID-19, caused by the SARS-CoV-2 virus, had a profound impact on Ethiopia's pharmaceutical supply chain. The pandemic disrupted global manufacturing and supply, leading to shortages of essential medicines and raw materials, Ethiopia, like many other countries, relies heavily on imported pharmaceutical products. With the pandemic causing lockdowns and restrictions in manufacturing hubs like India and China, the supply of essential medicines and raw materials to Ethiopia was significantly disrupted. The impact of COVID-19 on sales was significant. After the announcement of COVID-19 in Ethiopia, the central EPSS pushed a six-month supply of products to hubs to prevent stock outs. This push was from March to August 2020. During this period, if products were not stocked out, sales continued to hubs and health facilities.

- Physical inventory: refers to the process of manually counting and verifying the actual quantities of pharmaceutical products available in stock at a specific location. The process is critical for maintaining the accuracy of inventory records, ensuring that medicines are available when needed, and preventing stockouts or overstocking.  Physical inventory counting periods also influenced sales. Stores closed during these periods, halting transactions. We observed increased sales before inventory counting periods, as hubs and facilities stocked up. July and August were identified as physical counting periods each year.

- Internal displacement: Occurs when people are forced to flee their homes but remain within their country's borders, typically due to conflicts, violence, or natural disasters. Conflicts, whether ethnic, political, or resource-driven, can lead to widespread instability, forcing large populations to move and creating humanitarian crises. Conflicts often lead to the destruction of infrastructure, including roads, bridges, and warehouses, making it difficult to transport pharmaceutical products to affected areas. This disruption can cause significant delays or prevent the delivery of essential medicines altogether. Internal displacements and conflicts in Ethiopia over the past five years also affected sales and distribution. We traced periods of instability and included them as predictors. The identified conflict periods were September 2017 to March 2018, February 2019 to June 2019, and November 2020 to July 2022. 

- Malaria seasonality: Refers to the predictable patterns and fluctuations in malaria incidence throughout the year, typically influenced by climate and environmental conditions. In many regions, malaria transmission peaks during and shortly after the rainy season, when conditions such as stagnant water pools create ideal breeding sites for the Anopheles mosquitoes that transmit the disease. Conversely, malaria cases often decline during the dry season when mosquito breeding sites are reduced. During peak malaria seasons, there is a significant surge in the demand for antimalarial drugs and other related treatments. Malaria seasonality was another significant predictor. Certain pharmaceuticals, like Artemether + Lumefanthrine and Rapid Diagnostic Test kits, were affected by malaria outbreaks. We identified epidemic periods affecting sales: September to December 2017, March to May 2018, September to December 2018, March to May 2019, September to December 2019, March to May 2020, September to December 2020, March to May 2021, September to December 2021, and March to May 2022.



## Forecasting models

**Multiple Linear Regression (MLR)**:

**ARIMA and ARIMAX (LSTM)**:

**Long Short-Term Memory neural network (LSTM)**: The LSTM model is a specialised form of recurrent neural network (RNN) used to model sequential data by capturing long-term dependencies \[\@<https://doi.org/10.1162/neco.1997.9.8.1735>\]. Unlike traditional RNNs, LSTMs can learn to retain information for longer time periods due to their unique architecture, which consists of several gates that control the flow of information. This makes LSTMs particularly effective for time series forecasting. In our implementation, we used a sequential model architecture, comprising one LSTM layer with 50 units, followed by dense layers, with the final output being a single linear unit. The Adam optimizer was employed to minimize the mean squared error, and the model was trained for 100 epochs using the keras_model_sequential() function from the Keras package in R.

**TimeGPT**: TimeGPT is the first pre-trained foundational model designed specifically for time series forecasting, developed by Nixtla [@[https://doi.org/10.48550/arXiv.2310.0358910.03589](https://doi.org/10.48550/arXiv.2310.03589)]. It utilizes a transformer-based architecture with an encoder-decoder configuration but differs from other models in that it is not based on large language models. Instead, it is built from the ground up to handle time series data. TimeGPT was trained on more than 100 billion data points, drawing from publicly available time series across various sectors, such as retail, healthcare, transportation, demographics, energy, banking, and web traffic. This wide range of data sources, each with unique temporal patterns, enables the model to manage diverse time series characteristics effectively. Furthermore, TimeGPT supports the inclusion of external regressors in its forecasts and can generate quantile forecasts, providing reliable uncertainty estimation.

## Performance evaluation

To evaluate the performance of the various forecasting approaches, we split the data into a series of ten training and test sets. We use a time series cross-validation approach [@hyndman2021forecasting], with a forecast horizon of 84 days, and each training set expanding in 42-day steps. The first training set uses all data up to 2018-04-25, and the first test set uses the 84 days beginning 2018-04-26. The second training set uses all data up to 2018-06-06, with the second test set using the following 84 days. The largest training set ends on 2019-05-09, with the test set ending on 2019-07-31. Model development and hyper-parameter tuning is performed using the training data and the errors are assess using the corresponding test set. While we compute forecast errors for the entire 12 weeks, we are most interested in the last 42 days of each test set, because that corresponds to how forecasts are generated for planning in practice. Forecasting performance is evaluated using both point and probabilistic error measures.

The error metrics provided below consider a forecasting horizon denoted by $j$, representing the number of time periods ahead we are predicting. In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.

Point forecast accuracy is measured via the Mean Squared Scaled Error (MSSE) and the Mean Absolute Scaled Error (MASE). The Mean Absolute Scaled Error (MASE) [@HK06; @hyndman2021forecasting] is calculated as:
$$
  \text{MASE} = \text{mean}(|q_{j}|),
$$
where
$$
  q_{j} = \frac{ e_{j}}
 {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|},
$$
and $e_{j}$ is the point forecast error for forecast horizon $j$, $m = 7$ (as we have daily seasonal series), $y_t$ is the observation for period $t$, and $T$ is the sample size (the number of observations used for training the forecasting model). The denominator is the mean absolute error of the seasonal naive method in the fitting sample of $T$ observations and is used to scale the error. Smaller MASE values suggest more accurate forecasts. Note that the measure is scale-independent, thus allowing us to average the results across series.

A related measure is MSSE [@hyndman2021forecasting;@makridakis2022m5], which uses squared errors rather than absolute errors:
$$
  \text{MSSE} = \text{mean}(q_{j}^2),
$$ {#eq-RMSSE}
where,
$$
  q^2_{j} = \frac{ e^2_{j}}
 {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
$$
Again, this is scale-independent, and smaller MSSE values suggest more accurate forecasts.

Using scale-independent measures, such as MASE and MSSE, enables more appropriate comparisons between time series at different levels and scales, as these measures are not influenced by the magnitude of the data. This is of particular importance in our study, as we work with time series at various levels of hierarchy, with varying scales, resulting in different magnitudes of error. By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.

To measure the forecast distribution accuracy, we calculate the Continuous Rank Probability Score [@gneiting2014probabilistic;@hyndman2021forecasting]. It rewards sharpness and penalizes miscalibration, so it measures overall performance of the forecast distribution.
$$
  \text{CRPS} = \text{mean}(p_j),
$$ {#eq-CRPS}
where
$$
  p_j = \int_{-\infty}^{\infty} \left(G_j(x) - F_j(x)\right)^2dx,
$$
where $G_j(x)$ is the forecasted probability distribution function for forecast horizon $j$, and $F_j(x)$ is the true probability distribution function for the same period.

Calibration refers to the statistical consistency between the distributional forecasts and the observations. It measures how well the predicted probabilities match the observations. On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast. A model is well-calibrated if the predicted probabilities match the distribution of the observations, and it is sharp if it is confident in its predictions. The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated. Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.

CRPS can be considered an average of all possible Winkler scores [@winkler1972decision;@hyndman2021forecasting, Section 5.9] or percentile scores [@hyndman2021forecasting, Section 5.9], and thus provides an evaluation of all possible prediction intervals or quantiles. A specific prediction interval could be evaluated using a Winkler score. Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g. 95\%) quantiles. In such cases, a percentile score becomes useful in meeting this specific requirement.
